{\bf Landslide: Systematic Dynamic Race Detection in Kernel-space}

### 2012-05-09 defence/defence.tex
update defence for carlo

	\begin{itemize}
		\item On vanish, child's TCB is freed
		\item Forking thread does use-after-free
		\item Might return garbage instead of TID
	\end{itemize}
\end{frame}

@@

\begin{frame}{Finding Races in Kernel-Space} % slide: 15-410 6-week project why we need it in kernels
	\textbf{15-410:} Operating System Design and Implementation
	\begin{itemize}
		\item Project 3: students write a kernel in 6 weeks.
		\item ``Pebbles'' is a UNIX-like, preemptible kernel specification.
	\end{itemize}
	\linegap

	Previous work in kernel-space has not used systematic exploration.
	\begin{itemize}
		\item Data race detection\related{Erickson '10}
		\item Failure resiliency\related{Kadav '09}
	\end{itemize}
\end{frame}

@@

\begin{frame}{The Kernel as One Program} % slide: the kernel is too big to test all at once
	\begin{itemize}
		\item ``Everything interleaves with everything else''?
		\item What is a ``bug'', anyway?\related{Baumann '09}
	\end{itemize}
	\begin{center}
	\includegraphics[width=\textwidth]{pebbles.png}
	\end{center}
	% Say: "we don't want to have to explore every possible interaction between each of these components"
\end{frame}

@@

\begin{frame}{Outline}
	\done{Motivation: Concurrency debugging}
	\begin{itemize}
		\item \done{Systematic exploration versus stress testing}
		\item \done{Challenges of kernel-space}
	\end{itemize}
	\linegap

	{\bf Tool: Landslide}
	\begin{itemize}
		\item {\bf Design and interface}
		\item {\bf Addressing challenges}
	\end{itemize}
	\linegap

	Evaluation: Finding Races
	\begin{itemize}
		\item Student user study
		\item Case study
	\end{itemize}
	\linegap

	Future Work and Conclusion
\end{frame}

@@

\begin{frame}{Using Landslide}
	\begin{center}
	\includegraphics[width=0.9\textwidth]{tell_landslide.png}
	\end{center}
	User must tell Landslide when certain kernel events happen:
	\begin{itemize}
		\item When do threads become runnable / descheduled?
		\item When does the scheduler switch threads?
	\end{itemize}
	\linegap
	Some complicated conditions require in-Landslide instrumentation.
	\begin{itemize}
		\item Scheduler design details
	\end{itemize}
\end{frame}

\begin{frame}{Decision Trace}
	% TODO make the stack trace coloured
	\includegraphics[width=\textwidth]{found_a_bug.png}
\end{frame}

@@

\begin{frame}{State Space Reduction}
	State spaces grow exponentially.
	\begin{itemize}
		\item Fortunately, some sequences result in identical states.
		\item Dynamic Partial Order Reduction\related{Flanagan '05}
		\begin{itemize}
			\item Requires ``memory independence relation'' between transitions.
		\end{itemize}
	\end{itemize}
\end{frame}
\begin{frame}{State Space Reduction}
	\begin{center}
	\includegraphics[width=\textwidth]{undiamond1.png}
	\end{center}
\end{frame}

\begin{frame}{Shared Memory Conflicts} % slide: ignoring shared memory accesses - venn diagram
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict0.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict1.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict2.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\textbf{Problem:} Many conflicts we don't care about
	\begin{itemize}
		\item Every transition accesses scheduler state.
		\item Many transitions access ``unrelated'' global objects.
	\end{itemize}
	\linegap

	{\bf Solution:} Only consider ``relevant'' memory conflicts.
	\begin{itemize}
		\item Ignore scheduler, global objects ``known to be correct''
		\begin{itemize}
			\item Sacrifice ability to test these to efficiently test everything else.
		\end{itemize}
		\item User must specify what to ignore.
	\end{itemize}
\end{frame}

@@

	% TODO: does this slide want to be multicolumn, with an infinite loop diagram?
	Accurate conditions
	\begin{itemize}
		\item Kernel panic
		\item Use-after-free
		\item Deadlock
	\end{itemize}
	Uncertain conditions
	\begin{itemize}
		\item Memory leak
		\item Infinite loop
		\begin{itemize}
			\item Use structure of execution tree to judge stuckness.
		\end{itemize}
	\end{itemize}
\end{frame}
% slide: other_prob: decide on mutex_lock as opposed to everywhere, but...
% slide: manual shaping - have *user* specify what to focus on (automating intuition)
\begin{frame}{Limiting Decision Points}
\end{frame}
% slide: exciting teaser "i think this is not too far off!"
\begin{frame}{Teaser}
\end{frame}

@@

\begin{frame}{Outline}
	\done{Motivation: Concurrency debugging}
	\begin{itemize}
		\item \done{Systematic exploration versus stress testing}
		\item \done{Challenges of kernel-space}
	\end{itemize}
	\linegap

	\done{Tool: Landslide}
	\begin{itemize}
		\item \done{Design and interface}
		\item \done{Addressing challenges}
	\end{itemize}
	\linegap

	{\bf Evaluation: Finding Races}
	\begin{itemize}
		\item {\bf Student user study}
		\item {\bf Case study}
	\end{itemize}
	\linegap

	Future Work and Conclusion
\end{frame}

@@

	Five groups participated; four got it to work.
	\begin{itemize}
		\item Average instrumentation time 100 minutes (50 to 150)
		\item Average customisation time 25 minutes (0 to 60)
	\end{itemize}
\end{frame}

@@

\begin{frame}{Outline}
	\done{Motivation: Concurrency debugging}
	\begin{itemize}
		\item \done{Systematic exploration versus stress testing}
		\item \done{Challenges of kernel-space}
	\end{itemize}
	\linegap

	\done{Tool: Landslide}
	\begin{itemize}
		\item \done{Design and interface}
		\item \done{Addressing challenges}
	\end{itemize}
	\linegap

	\done{Evaluation: Finding Races}
	\begin{itemize}
		\item \done{Student user study}
		\item \done{Case study}
	\end{itemize}
	\linegap

@@

	After Pebbles\dots
	\begin{itemize}
		\item Potential application to production kernels
		\begin{itemize}
			\item General purpose OSes, device drivers, embedded kernels
			% Be ready to field questions about SMP.
		\end{itemize}
	\end{itemize}
	\linegap

@@

%\subsection{Summary}
\begin{frame}{Summary}
	\textbf{Systematic exploration in kernel-space}
	\begin{itemize}
		\item Use internal kernel abstractions to understand concurrency behaviour.
		\item Relying on user's knowledge makes testing easier.
	\end{itemize}
	\linegap
	\textbf{Landslide}
	\begin{itemize}
		\item After polishing, a useful tool for 15-410.
		\item A first step towards sophisticated kernel debugging techniques.
	\end{itemize}

>>>>

	\begin{itemize}
		\item On exit, child's TCB is freed
		\item Forking thread does use-after-free
		\item Might return garbage instead of TID
	\end{itemize}
\end{frame}

@@

\begin{frame}{Finding Races in Kernel-Space} % slide: 15-410 6-week project why we need it in kernels
	Previous work in kernel-space has not used systematic exploration.
	\linegap

	\textbf{15-410:} Operating System Design and Implementation
	\begin{itemize}
		\item Project 3: students write a kernel in 6 weeks.
		\item ``Pebbles'' is a UNIX-like, preemptible kernel specification.
	\end{itemize}
	%\begin{itemize}
	%	\item Data race detection\related{Erickson '10}
	%	\item Failure resiliency\related{Kadav '09}
	%\end{itemize}
\end{frame}

@@

\begin{frame}{The Kernel as One Program} % slide: the kernel is too big to test all at once
	\begin{center}
	\includegraphics[width=\textwidth]{pebbles.png}
	\end{center}
\end{frame}
\begin{frame}{The Kernel as One Program} % slide: the kernel is too big to test all at once
	\begin{itemize}
		\item ``Everything interleaves with everything else''?
		\item What is a ``bug'', anyway?\related{Baumann '09}
	\end{itemize}
	% Say: "we don't want to have to explore every possible interaction between each of these components"
\end{frame}

@@

%\begin{frame}{Outline}
%	\done{Motivation: Concurrency debugging}
%	\begin{itemize}
%		\item \done{Systematic exploration versus stress testing}
%		\item \done{Challenges of kernel-space}
%	\end{itemize}
%	\linegap
%
%	{\bf Tool: Landslide}
%	\begin{itemize}
%		\item {\bf Design and interface}
%		\item {\bf Addressing challenges}
%	\end{itemize}
%	\linegap
%
%	Evaluation: Finding Races
%	\begin{itemize}
%		\item Student user study
%		\item Case study
%	\end{itemize}
%	\linegap
%
%	Future Work and Conclusion
%\end{frame}

@@

\begin{frame}{Using Landslide}
	\begin{center}
	\includegraphics[width=0.9\textwidth]{tell_landslide.png}
	\end{center}
	User must tell Landslide when certain kernel events happen:
	\begin{itemize}
		\item When do threads become runnable / descheduled?
		\item When does the scheduler switch threads?
	\end{itemize}
	\linegap
	% TODO
	Some complicated conditions require in-Landslide instrumentation.
	\begin{itemize}
		\item Scheduler design details
	\end{itemize}
\end{frame}

\begin{frame}{Decision Trace}
	% FIXME make the stack trace coloured
	\includegraphics[width=\textwidth]{found_a_bug.png}
\end{frame}

@@

\begin{frame}{State Space Reduction}
	\begin{center}
	\includegraphics[width=\textwidth]{undiamond1.png}
	\end{center}
\end{frame}
\begin{frame}{State Space Reduction}
	State spaces grow exponentially.
	\begin{itemize}
		\item Fortunately, some sequences result in identical states.
		\item Dynamic Partial Order Reduction\related{Flanagan '05}
		\begin{itemize}
			\item Requires ``memory independence relation'' between transitions.
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Shared Memory Conflicts} % slide: ignoring shared memory accesses - venn diagram
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict0.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict1.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict2.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	%\textbf{Problem:} Many conflicts we don't care about
	%\begin{itemize}
	%	\item Every transition accesses scheduler state.
	%	\item Many transitions access ``unrelated'' global objects.
	%\end{itemize}
	%\linegap
	\textbf{Solution:} Only consider ``relevant'' memory conflicts.
	\begin{itemize}
		\item Ignore scheduler, global objects assumed to be correct
		\begin{itemize}
			\item Sacrifice ability to test these to efficiently test everything else.
		\end{itemize}
		\item User must specify what to ignore.
	\end{itemize}
\end{frame}

@@

	% TODO: does this slide want to be multicolumn, with an infinite loop diagram?
	Definite bugs
	\begin{itemize}
		\item Kernel panics / assertion failures
		\item Use-after-free
		\item Deadlock
	\end{itemize}
	Probable bugs
	\begin{itemize}
		\item Infinite loop
		\begin{itemize}
			\item Use structure of execution tree to judge stuckness.
		\end{itemize}
	\end{itemize}
\end{frame}
% slide: other_prob: decide on mutex_lock as opposed to everywhere, but...
% slide: manual shaping - have *user* specify what to focus on (automating intuition)
\begin{frame}{Limiting Decision Points}
\end{frame}
% slide: exciting teaser "i think this is not too far off!"
\begin{frame}{Teaser}
\end{frame}

@@

%\begin{frame}{Outline}
%	\done{Motivation: Concurrency debugging}
%	\begin{itemize}
%		\item \done{Systematic exploration versus stress testing}
%		\item \done{Challenges of kernel-space}
%	\end{itemize}
%	\linegap
%
%	\done{Tool: Landslide}
%	\begin{itemize}
%		\item \done{Design and interface}
%		\item \done{Addressing challenges}
%	\end{itemize}
%	\linegap
%
%	{\bf Evaluation: Finding Races}
%	\begin{itemize}
%		\item {\bf Student user study}
%		\item {\bf Case study}
%	\end{itemize}
%	\linegap
%
%	Future Work and Conclusion
%\end{frame}

@@

	Five groups participated; four got it to work.
	\begin{itemize}
		\item Average instrumentation time 100 minutes (50 to 150)
		%\item Average customisation time 25 minutes (0 to 60)
	\end{itemize}
\end{frame}

@@

%\begin{frame}{Outline}
%	\done{Motivation: Concurrency debugging}
%	\begin{itemize}
%		\item \done{Systematic exploration versus stress testing}
%		\item \done{Challenges of kernel-space}
%	\end{itemize}
%	\linegap
%
%	\done{Tool: Landslide}
%	\begin{itemize}
%		\item \done{Design and interface}
%		\item \done{Addressing challenges}
%	\end{itemize}
%	\linegap
%
%	\done{Evaluation: Finding Races}
%	\begin{itemize}
%		\item \done{Student user study}
%		\item \done{Case study}
%	\end{itemize}
%	\linegap
%
%	{\bf Future Work and Conclusion}
%\end{frame}

@@

	After Pebbles\dots
	\begin{itemize}
		\item Potential application to production kernels
		\begin{itemize}
			\item General purpose OSes, device drivers
			% Be ready to field questions about SMP.
		\end{itemize}
	\end{itemize}
	\linegap

@@

%\subsection{Summary}
\begin{frame}{Summary}
	\textbf{Systematic exploration in kernel-space}
	\begin{itemize}
		\item Use internal kernel abstractions to understand concurrency behaviour.
		\item Relying on user's knowledge makes testing easier.
	\end{itemize}
	\linegap
	\textbf{Landslide}
	\begin{itemize}
		\item Has potential to be a useful tool for 15-410.
		\item A first step towards sophisticated kernel debugging techniques.
	\end{itemize}
